Dynamic capabilities and competitive advantage

### 2019-09-26 article4_tommylars.rnw
Before write up

\documentclass[review,fleqn]{elsarticle}
\usepackage[textsize=tiny,colorinlistoftodos]{todonotes}
\usepackage{subfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{filecontents}
\input{preamble.tex}
%%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\setlength{\mathindent}{0pt} %% NB MAYBEREMOVE BEFORE SUBMISSION

@@

\section{Introduction}
In a recent "state of the art" literature review on dynamic capabilities
\cite{Schilke2018} state that much of the attention and interest surrounding the theory is
that it offer a route to competitive advantage which they argue is "the virtual Holy Grail
of strategic management" (p. 393). And indeed a considerable body of the extant literature
concerns itself with exactly this path to competitive advantage (refs from
shcilke). Still, a discourse regarding the contingencies of the dynamic
capabilities-competitive advantage (DC-CA) nexus (?) has panned out in the literature,
most notably coined "the elephant in the room" \citep{Peteraf2013}: Under which conditions
will dynamic capabilities lead to competitive advantage?

@@

\text{red}{CONTINUE WITH THE OTHER KEY CONSTRUCTS – INPUT FROM TOMMY}

@@

We add a set of control variables expected to explain parts of the variation in our focal
constructs. From the survey we add firm size, firm age, the level of dynamism facing the
firm. Size is the reported number of employees in the firm at $T_2$, and age is the number of
years since establishment. Dynamism is captured by asking a number of questions about the
competitive environment facing the firm. A full list of items and their internal validity
(0.75) is available in appendix ##.

@@

\begin{table}[t]
\begin{singlespace}
  \caption{\small Descriptive statistics and correlations for variables in this paper}
  \label{tab:desc}

<<resultscorr, results="asis",echo=FALSE >>=

print(xtable(art2cor ,size="\\small",caption="Correlation matrix \\label{tab:cor}", align = c("l", rep("p{0.3in}",9)),table.placement="h", size = "\\small", tabular.environment="longtable"), size="scriptsize", label='tab:cor', caption.placement="top", floating=FALSE,latex.environments=NULL)

@
\vspace{1ex}

     \raggedright {\scriptsize Notes: Pairwise correlations between variables in the study
    excluding controls}
\end{singlespace}
\end{table}

\begin{table*}[t]
  \centering

\caption{Regression -How are MDC influencing evolution of lower level capabilities }

<<result1, results="asis", echo=FALSE >>=
stargazer(mainres,
          add.lines =list(c("Adjusted $R^2$",r1,r2,r3,r4,r5,r6)),
          single.row=FALSE, column.sep.width = "1pt", omit.stat=c("LL","ser","f","aic","rsq","adj.rsq"),
          no.space=FALSE,
          dep.var.labels=c("OR_{T_2}", "DC_{T_2}","CA_{T_2}","DL"),
        #  column.labels=c("HECKMAN","OLS","HECKMAN","OLS","HECKMAN","OLS"),
          model.names=FALSE,
          object.names=FALSE,

          ### order=c(1,3,2),

           ## covariate.labels= c("$OR_{T_1}$",
           ##                "$\\Delta OR$",
           ##                "$CA_{T_1}$"
           ##                "$DL$",
           ##                "$DC_{T_1}$",
           ##                "$\\Delta DC$",
           ##                "Firm size",
           ##                "Firm age",
           ##                "Environmental Dynamism",
           ##                "Firm profit margin",
           ##                "Firm assets"
           ##          ),

          title= "Regression results for model C", font.size="scriptsize",float=FALSE
          )

@
\label{tab:reg}
\end{table*}

\begin{figure}[t]

\begin{minipage}[t]{.5\textwidth}

  \resizebox{7cm}{!}{\input{./figure/medanalysis1.tex}}
  \captionof{figure}{Mediation analysis 1}
  \label{fig:fig3}
\end{minipage}%
\begin{minipage}[t]{.5\textwidth}
 \resizebox{7cm}{!}{\input{./figure/medanalysis2.tex}}
  \captionof{figure}{Mediation analysis 2}
  \label{fig:fig4}
\end{minipage}
\end{figure}

Table \ref{tab:reg} shows a summary of the regression models proposed in the previous
chapter. Models 1,2 and 3 are designed to test DL as \emph{antecedent} for $OR$ and
$DC$. Consider model 1 we observe that $DL$ is influencing the level of $OR$ within the
firm ($\beta_2 =$ \Sexpr{mod1_dl}) supporting H1. However, the effect disappears when we
add $DC$ in model 2 suggesting a mediating effect. In other words we see that firms with
initial high $DC$($\beta_3=$ \Sexpr{mod2_dc0}) and increasing $DC$ over time ($\beta_4=$
\Sexpr{mod2_ddc}) also have better $OR$. To control for the mediating effect suggested by
the change in coefficients for $DL$ between models 1 and 2, we performed a causal
mediation analysis as reported in figure \ref{fig:fig3}. This analysis confirms the
indications from models 1 and 2 showing that the direct effect from $DL$ on $OR$ is
insignificant and small (ADE = \Sexpr{medmod1_ade}). Meanwhile, the indirect effect
running through changes in $DC$ over time, is significant yielding a positive and
significant indirect effect of \Sexpr{medmod1_acme}. The proportion of the effect that is
mediated amounts to \Sexpr{medmod1_shmed}. In sum these results suggest that H1 is
unsupported, and that the bulk of the effect is indirect running through $DC$, in line with
the expectations in H3. Similarly, we see from model 3 that $DC$ is indeed influenced by $DL$
above and beyond its own evolutionary trajectory. This is in line with our expectations
and lend support to H4.

To evaluate H5 proposing a feedback effect from $DC$ to $DL$, we turn to model 6 in table
\ref{tab:reg}. Controlled for the usual firm level controls we find a positive effect
($\xi_1=$ \Sexpr{mod6_dc0}) from $DC_{T_1}$ on $DL$ measured at $T_1$. This supports H5
suggesting that firms with more $DC$ is more likely to generate opportunities where it
becomes more prevalent that learning will need to take place. One interesting finding is
also that this effect environmental dynamism decreases firms level of $DL$. This is indeed
puzzling and we will look in to this in a post-hoc analysis.\todo{When adding interactions
  we see that high environmental dynamism in conjunction with DC is positive for DL
  meaning that DC will indeed work to make firms more aware of the benefits of DL when
  environments are changing} It is also worth noting that models 1,2 and 3 have positive
and significant $\rho$ from the Heckman selection model indicating that selection is
leading to biased estimates. By running a Heckman correction we avoid this potential
problem.

Models 4 and 5 in table \ref{tab:reg} deals with the \emph{outcome} of DC. H6a and H6b
both considers effects of $DC$ on the competitive advantage of the firm $CA$. $CA$ is
influenced by the level of $DC$ of the firm indicating that dynamic capabilities are a
source of competitive advantage in line with our expectations. Furthermore, considering
model 5 we see that $OR$ also has a significant positive effect on $CA$ indicating that
operating routines can also be a source of advantage for the firm. Moreover, the
coefficient of $\Delta DC$ decline from \Sexpr{mod4_ddc} to \Sexpr{mod5_ddc} when we add
$OR$. This suggests that some mediation is at work, i.e. that changes in $DC$ leads to
changes in changes in $OR$ leading to $CA$. This indirect path is in line with the
expectations from our hypotheses and as described in figure \ref{fig:fig2}. When formally
testing this mediation running a causal mediation analysis (see figure \ref{fig:fig4}), we
find added support that both a \emph{direct-} and an \emph{indirect} path exist to explain
the relationship between $DC$ and $CA$ (ADE = \Sexpr{medmod5_ade}, ACME =
\Sexpr{medmod5_acme}) and a total mediated share of \Sexpr{medmod5_shmed}.

>>>>

\documentclass[review,fleqn]{elsarticle}
\usepackage[textsize=tiny,colorinlistoftodos]{todonotes}
\usepackage{subfig}
\usepackage{subcaption}

\input{preamble.tex}
%%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\setlength{\mathindent}{0pt} %% NB MAYBEREMOVE BEFORE SUBMISSION

@@

\section{Introduction}

\Sexpr{resest[1,2]}
In a recent "state of the art" literature review on dynamic capabilities
\cite{Schilke2018} state that much of the attention and interest surrounding the theory is
that it offer a route to competitive advantage which they argue is "the virtual Holy Grail
of strategic management" (p. 393). And indeed a considerable body of the extant literature
concerns itself with exactly this path to competitive advantage (refs from
shcilke). Still, a discourse regarding the contingencies of the dynamic
capabilities-competitive advantage (DC-CA) nexus (?) has panned out in the literature,
most notably coined "the elephant in the room" \citep{Peteraf2013}: Under which conditions
will dynamic capabilities lead to competitive advantage?

@@

<<result1, results="asis", echo=FALSE >>=

ht

@

@@

We add a set of control variables expected to explain parts of the variation in our focal
constructs. From the survey we add firm size, firm age, the level of dynamism facing the
firm. Size is the reported number of employees in the firm at $T_2$, and age is the number of
years since establishment. Dynamism is captured by asking a number of questions about the
competitive environment facing the firm. A full list of items and their internal validity
(0.75) is available in appendix 1111.

@@

\begin{figure}[ht] \label{ fig7} 
  \begin{minipage}[b]{0.5\linewidth}
    \includegraphics[width=.99\linewidth]{full.pdf} 
    \caption{Full model} 
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \includegraphics[width=.99\linewidth]{ma.pdf} 
    \caption{Routine mediation} 
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \includegraphics[width=.99\linewidth]{or.pdf} 
    \caption{Process mediation} 
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.5\linewidth}
    \includegraphics[width=.99\linewidth]{dir.pdf} 
    \caption{Direct model} 
  \end{minipage} 
\end{figure}

\begin{figure}
     \centering
  \caption{Structural model of full estimation - PLS-SEM}
  \input{./figure/fullmod.tex}
\end{figure}

\begin{table}
\begin{center}
\script size
\begin{tabular}{lcccc}

\hline \\[-2pt]
 & \multicolumn{4}{c}{{\bf Models:}} \\  \cline{2-5} \\[-3pt] 
 & \emph{Direct} & \emph{Process} & \emph{Routines} & \emph{Full}\\
 & (1) & (2) & (3) & (4)\\
\hline
 &  &  &  & \\
\(CA_{t-1} \rightarrow CA\) & \Sexpr{plmest[1,2]} & \Sexpr{plmest[1,3]} & \Sexpr{plmest[1,4]} & \Sexpr{plmest[1,5]}\\
 & (\Sexpr{plmt[1,2]}) & (\Sexpr{plmt[1,3]}) & (\Sexpr{plmt[1,4]}) & (\Sexpr{plmt[1,5]})\\[3pt]
 &  &  &  & \\
\(DC_P \rightarrow CA\) & \Sexpr{plmest[2,2]} & \Sexpr{plmest[2,3]} & \Sexpr{plmest[2,4]} & \Sexpr{plmest[2,5]}\\
 & (\Sexpr{plmt[2,2]}) & (\Sexpr{plmt[2,3]}) & (\Sexpr{plmt[2,4]}) & (\Sexpr{plmt[2,5]})\\[4pt]
\(DC_X \rightarrow CA\) & \Sexpr{plmest[3,2]} & \Sexpr{plmest[3,3]} & \Sexpr{plmest[3,4]} & \Sexpr{plmest[3,5]}\\
 & (\Sexpr{plmt[3,2]}) & (\Sexpr{plmt[3,3]}) & (\Sexpr{plmt[3,4]}) & (\Sexpr{plmt[3,5]})\\[4pt]
 &  &  &  & \\
\(DC_P \rightarrow P\) &  & \Sexpr{plmest[4,3]} &  & \Sexpr{plmest[4,5]}\\
 &  & (\Sexpr{plmt[4,3]}) &  & (\Sexpr{plmt[4,5]})\\[4pt]
\(DC_P \rightarrow R\) &  &  & \Sexpr{plmest[5,4]} & \Sexpr{plmest[5,5]}\\
 &  &  & (\Sexpr{plmt[5,4]}) & (\Sexpr{plmt[5,5]})\\[4pt]
\(DC_X \rightarrow P\) &  & \Sexpr{plmest[6,3]} &  & \Sexpr{plmest[6,5]}\\
 &  & (\Sexpr{plmt[6,3]}) &  & (\Sexpr{plmt[6,5]})\\[4pt]
\(DC_X \rightarrow R\) &  &  & \Sexpr{plmest[7,4]} & \Sexpr{plmest[7,5]}\\
 &  &  & (\Sexpr{plmt[7,4]}) & (\Sexpr{plmt[7,5]})\\[4pt]
 &  &  &  & \\
\(P \rightarrow CA\) &  & \Sexpr{plmest[8,3]} &  & \Sexpr{plmest[8,5]}\\
 &  & (\Sexpr{plmt[8,3]}) &  & (\Sexpr{plmt[8,5]})\\[4pt]
\(R \rightarrow CA\) &  &  & \Sexpr{plmest[9,4]} & \Sexpr{plmest[9,5]}\\
 &  &  & (\Sexpr{plmt[9,4]}) & (\Sexpr{plmt[9,5]})\\[4pt]
 &  &  &  & \\
\hline \\[-3pt] 
Goodness of fit & \Sexpr{gofdir} & \Sexpr{gofor} & \Sexpr{gofma} & \Sexpr{goffull}\\
Number of observations & \Sexpr{N} & \Sexpr{N} & \Sexpr{N} & \Sexpr{N}\\
  Firm level controls & YES & YES & YES & YES\\[3pt]
\hline
 &  &  &  & \\
\end{tabular}
\end{center}
 \label{tab:reg}
\end{table}

Table \ref{tab:reg} shows a summary of the regression models proposed in the previous
chapter. Models 1,2 and 3 are designed to test DL as \emph{antecedent} for $OR$ and
$DC$. Consider model 1 we observe that $DL$ is influencing the level of $OR$ within the
firm ($\beta_2 =$r) supporting H1. However, the effect disappears when we
add $DC$ in model 2 suggesting a mediating effect. In other words we see that firms with
initial high $DC$($\beta_3=$ \ {mod2_dc0}) and increasing $DC$ over time ($\beta_4=$
\ {mod2_ddc}) also have better $OR$. To control for the mediating effect suggested by
the change in coefficients for $DL$ between models 1 and 2, we performed a causal
mediation analysis as reported in figure \ref{fig:fig3}. This analysis confirms the
indications from models 1 and 2 showing that the direct effect from $DL$ on $OR$ is
insignificant and small (ADE = \ {medmod1_ade}). Meanwhile, the indirect effect
running through changes in $DC$ over time, is significant yielding a positive and
significant indirect effect of \ {medmod1_acme}. The proportion of the effect that is
mediated amounts to \ {medmod1_shmed}. In sum these results suggest that H1 is
unsupported, and that the bulk of the effect is indirect running through $DC$, in line with
the expectations in H3. Similarly, we see from model 3 that $DC$ is indeed influenced by $DL$
above and beyond its own evolutionary trajectory. This is in line with our expectations
and lend support to H4.

To evaluate H5 proposing a feedback effect from $DC$ to $DL$, we turn to model 6 in table
\ref{tab:reg}. Controlled for the usual firm level controls we find a positive effect
($\xi_1=$ \ {mod6_dc0}) from $DC_{T_1}$ on $DL$ measured at $T_1$. This supports H5
suggesting that firms with more $DC$ is more likely to generate opportunities where it
becomes more prevalent that learning will need to take place. One interesting finding is
also that this effect environmental dynamism decreases firms level of $DL$. This is indeed
puzzling and we will look in to this in a post-hoc analysis.\todo{When adding interactions
  we see that high environmental dynamism in conjunction with DC is positive for DL
  meaning that DC will indeed work to make firms more aware of the benefits of DL when
  environments are changing} It is also worth noting that models 1,2 and 3 have positive
and significant $\rho$ from the Heckman selection model indicating that selection is
leading to biased estimates. By running a Heckman correction we avoid this potential
problem.

Models 4 and 5 in table \ref{tab:reg} deals with the \emph{outcome} of DC. H6a and H6b
both considers effects of $DC$ on the competitive advantage of the firm $CA$. $CA$ is
influenced by the level of $DC$ of the firm indicating that dynamic capabilities are a
source of competitive advantage in line with our expectations. Furthermore, considering
model 5 we see that $OR$ also has a significant positive effect on $CA$ indicating that
operating routines can also be a source of advantage for the firm. Moreover, the
coefficient of $\Delta DC$ decline from  {mod4_ddc} to  {mod5_ddc} when we add
$OR$. This suggests that some mediation is at work, i.e. that changes in $DC$ leads to
changes in changes in $OR$ leading to $CA$. This indirect path is in line with the
expectations from our hypotheses and as described in figure \ref{fig:fig2}. When formally
testing this mediation running a causal mediation analysis (see figure \ref{fig:fig4}), we
find added support that both a \emph{direct-} and an \emph{indirect} path exist to explain
the relationship between $DC$ and $CA$ (ADE = \ {medmod5_ade}, ACME =
\ {medmod5_acme}) and a total mediated share of \ {medmod5_shmed}.
